---
author: "smooshi"
date: "2018 M11 6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 3

##Task 1-2
```{r}
alc <- read.table("~/GitHub/IODS-project/data/alc.csv", header = TRUE, sep = ",")
names(alc)
```

This is a dataset made from two Portuguese school data attributes include descriptors of students (like age, sex, family size, profession of mom and dad, grades etc) and alcohol usage (Dalc = workday alcohol consumption, Walc = weekend alcohol consumption). Alc_use combines Dalc and Walk, and high_use is alcohol usage that's higher than 2 (1- very low, 5- very high). 

##Task 3-4

###Age

Legal drinking age in Portugal is 16 for beer/cider and 18 for hard liquor. I would hope that there are significantly more high alchohol users in students that are of legal drinking age.

```{r}
library(ggplot2)
alc$legal <- ifelse(alc$age >= 16, TRUE, FALSE)
ggplot(data = alc, aes(x=high_use))+geom_bar()+facet_wrap("legal")+geom_text(stat='count', aes(label=..count..))

#Students under legal drinking age that are high users
18/(63+18)

#Students over legal drinking age that are high users
96/(205+96)
```

Here groupings are based on is the student legal drinking age (TRUE/FALSE). The graph shows that there are much more responders that are of legal drinking age. Percentage of students that are high users and under the legal drinking age is 22% and number of students that are high users and over the legal drinking age is 32%. There's a noticable difference.

### Go out

26 goout - going out with friends (numeric: from 1 - very low to 5 - very high) 

Assumption: Kids that go out alot with their friends are more likely to drink than those who don't.

```{r}
ggplot(data = alc, aes(x=high_use))+geom_bar()+facet_wrap("goout")+geom_text(stat='count', aes(label=..count..))

ftable(table(alc$high_use, alc$goout))
```

Here it's very obvious, kids that go out a lot drink a lot more than kids that don't go out much.



### Grades

Assumption: Kids who get more than half the points in every grade (First period, second period, final grade) are less likely to drink a lot than kids than don't even manage half the points.

```{r}
alc$grades <- ifelse(alc$G1 >= 11 & alc$G2 >= 11 & alc$G3 >=11, TRUE, FALSE)
ggplot(data = alc, aes(x=high_use))+geom_bar()+facet_wrap("grades")+geom_text(stat='count', aes(label=..count..))

#false
66/(66+118)

#true
48/(48+150)
```

This looks like it might be significant, kids that get more than half the points are 24% high alcohol users and kids that don't are 35.9% high alchohol users.

###Studytime

Assumption: Kids who spend more time studying drink less.

```{r}
ggplot(data = alc, aes(x=high_use))+geom_bar()+facet_wrap("studytime")+geom_text(stat='count', aes(label=..count..))

ftable(table(alc$high_use, alc$studytime))

#1 < 2hours
42/(42+58)

#2 2 to 5 hours
60/(60+135)

#3 5 to 10 hours
8/(8+52)

#4 over 10 hours
4/(4+23)
```

Here it's pretty clear, kids that study more than 5 hours a week are much less likely to be drinkers (14.8%, 13.3%) and kids who study less are much more likely to be heavy drinkers ( 30.7% and 42%). Especially if the kids spend almost no time studying (<2).

##Task 5

###Age

```{r}
m <- glm(high_use ~ legal, data = alc, family = "binomial")
summary(m)
coef(m)
confint(m)
```

### Go out

```{r}
m <- glm(high_use ~ goout, data = alc, family = "binomial")
summary(m)
coef(m)
confint(m)
```

###Grades

```{r}
m <- glm(high_use ~ grades, data = alc, family = "binomial")
summary(m)
coef(m)
confint(m)
```

###Studytime

```{r}
m <- glm(high_use ~ studytime, data = alc, family = "binomial")
summary(m)
coef(m)
confint(m)
```