# R Exercise 2 Analysis

##Task 1

Data wrangling exersize is in the data folder.

```{r}
data <- read.table("~/GitHub/IODS-project/data/learning2014.csv", header = TRUE, sep = ",")
head(data)
names(data)
dim(data)
```

##Task 2
```{r}
library(ggplot2)
summary(data)
```
Summary of all variables. Dataset is about students (Age, Gender) and their attitudes towards learning (Attitude, Deep, Stra, Surf) and their combined scores (Points). I will use the graphical library ggplot to do graphics.

```{r}
#Pie chart of gender
ggplot(data, aes(x=factor(1), fill = factor(Gender)))+geom_bar(width=1)+coord_polar(theta="y")+theme_void()
```
There are 122 females and 61 males.

```{r}
#Histogram of Age
qplot(data$Age, geom="histogram")+labs(title="Histogram for Age") +
  labs(x="Age", y="Count")
```
Age distribution is slanted, there are a lot more people in their 20's than other people.

```{r}
#Boxplot of gender and points

ggplot(data, aes(x=Gender, y=Points, fill=Gender)) + 
  geom_boxplot()
```
It looks like Males have slightly more points by average than females.

Points and Attitude:

```{r}
#Linear regression of Attitude and Points
ggplot(data, aes(x = Attitude, y = Points)) + geom_point() + geom_smooth(method="lm")+labs(title="Linear regression of attitude and points")
```

Scatter plot matrix:
```{r}
library(GGally)
p <- ggpairs(data, mapping = aes(col=Gender, alpha=0.3), lower = list(combo = wrap("facethist", bins = 20)))
p
```

##Task 3 & 4

###Task 3
Target: Points

```{r}
linear <- lm(Points ~ Gender+Age+Attitude, data=data)
summary(linear)
```


This correlation is not significant for gender and age, but is for attitude. 

```{r}
summary(lm(Points~Attitude+Stra+Surf, data=data))
```

Linear regression is significant for Attitude and Stra, but not Surf

###Task 4

```{r}
summary(lm(Points~Attitude+Stra+Deep, data=data))
```
Attitude, Strategy, and Deep understanding during studying explains exam results the best. Here multiple linear regression and all subregressions are significant. 

R-squared is a statistical measure of how close the data are to the fitted regression line. Multiple R-squared is low and that says that these variables do not explain the whole variation in exam points.

```{r}
ggplot(data, aes(x=Points, y=Attitude+Stra+Deep))+geom_point()+geom_smooth(method="lm")+labs(title="Linear regression of  points and attitude+stra+deep")
```

In the picture there's slight positive correlation between Points and Attitude+Stra+Deep. While Points correlates with Attitude+Stra+Deep, those variables do not explain all variation in Points.

##Task 5

```{r}
linear <- lm(Points~Attitude+Stra+Deep, data=data)
par(mfrow = c(2,2))
plot(linear, which = c(1,2,5))
```
Q-Q Plot shows that the left tail of the distribution does not follow normalcy. It deviates a lot.
There is also high spread in residuals, but they look mostly grouped linearly.
There is one higher leverage outlier in the data.