---
title: "Chapter6"
author: "smooshi"
date: "2018 M12 9"
output: html_document
---

```{r setup6, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
```

#Chapter 6

```{r}
BPRS <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/BPRS.txt", sep=" ", header=T)
rats <- read.csv("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/rats.txt", sep="\t", header=T)

#Factorize
BPRS$treatment <- factor(BPRS$treatment)
BPRS$subject <- factor(BPRS$subject)
rats$Group <- factor(rats$Group)
rats$ID <- factor(rats$ID)

#Long
BPRSL <-  BPRS %>% gather(key = weeks, value = bprs, -treatment, -subject)
#Extract week/wd
BPRSL <-  BPRSL %>% mutate(week = as.integer(substr(weeks,5,5)))

ratsL <-  rats %>% gather(key = WD, value = Weight, -ID, -Group)  %>%
  mutate(Time = as.integer(substr(WD,3,4))) 


glimpse(ratsL)
glimpse(BPRSL)
```

##BPRS

### Data at first
```{r}
summary(BPRSL)

ggplot(BPRSL, aes(x = week, y = bprs, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(BPRSL$bprs), max(BPRSL$bprs)))
```


Data before stanardization.

### Normalized data 
```{r}
BPRSL <- BPRSL %>%
  group_by(week) %>%
  mutate(stdbprs = (bprs - mean(bprs))/sd(bprs) ) %>%
  ungroup()

# Glimpse the data
glimpse(BPRSL)

# Plot again with the standardised bprs
ggplot(BPRSL, aes(x = week, y = stdbprs, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  scale_y_continuous(name = "standardized bprs")
```

Data after standardization. The plot suggests the presence of some possible outliers and indicates again the general decline in BPRS values over the eight weeks of the study in both treatment groups.

### Mean response profiles:
```{r}
# Number of weeks, baseline (week 0) included
n <- BPRSL$week %>% unique() %>% length()

# Summary data with mean and standard error of bprs by treatment and week 
BPRSS <- BPRSL %>%
  group_by(treatment, week) %>%
  summarise( mean = mean(bprs), se = sd(bprs)/sqrt(n) ) %>%
  ungroup()

# Glimpse the data
glimpse(BPRSS)

# Plot the mean profiles
ggplot(BPRSS, aes(x = week, y = mean, linetype = treatment, shape = treatment)) +
  geom_line() +
  scale_linetype_manual(values = c(1,2)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se, linetype="1"), width=0.3) +
  theme(legend.position = c(0.8,0.8)) +
  scale_y_continuous(name = "mean(bprs) +/- se(bprs)")
```

We see BPRS values decrease in both treatment groups over time. However treatment 2 bumps in the end.

### Remove Outliers

```{r}
# Create a summary data by treatment and subject with mean as the summary variable (ignoring baseline week 0).
BPRSL8S <- BPRSL %>%
  filter(week > 0) %>%
  group_by(treatment, subject) %>%
  summarise( mean=mean(bprs) ) %>%
  ungroup()

# Glimpse the data
glimpse(BPRSL8S)

# Draw a boxplot of the mean versus treatment
ggplot(BPRSL8S, aes(x = treatment, y = mean)) +
  geom_boxplot() +
  stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(bprs), weeks 1-8")

# Create a new data by filtering the outlier and adjust the ggplot code the draw the plot again with the new data
BPRSL8S1 <- filter(BPRSL8S, mean < 60)
```

Outlier is seen as a black dot above boxplots. It is around ~70, but I removed everything over 60 (over the whiskers of the boxplots)

Plot after removal:

```{r}
ggplot(BPRSL8S1, aes(x = treatment, y = mean)) +
  geom_boxplot() +
  stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(bprs), weeks 1-8")
```

###Anova:

```{r}
# Perform a two-sample t-test
t.test(mean ~ treatment, data = BPRSL8S1, var.equal = TRUE)

# Add the baseline from the original data as a new variable to the summary data
BPRSL8S2 <- BPRSL8S %>%
  mutate(baseline = BPRS$week0)

# Fit the linear model with the mean as the response 
fit <- lm(mean ~ baseline+treatment, data = BPRSL8S2)

# Compute the analysis of variance table for the fitted model with anova()
anova(fit)
```

T-test shows no significant difference between treatments, but when baseline week0 is incorporated we see that baseline BPRS is strongly related to the BPRS values after treatment... But treatment is still not significant.

## Rats

```{r}
ggplot(ratsL, aes(x = Time, y = Weight, group = ID)) +
  geom_line(aes(linetype = Group)) + scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 10)) + scale_y_continuous(name = "Weight (grams)") + theme(legend.position= "top")
```


Rats in groups 2 and 3 look like they're gaining weight. Especially group 2? And also repeated measures of weight are not independant of each other.

###Linear regression

```{r}
# create a regression model RATS_reg
RATS_reg <- lm(Weight ~ Time+Group, data=ratsL)

# print out a summary of the model
summary(RATS_reg)

# access library lme4
library(lme4)

# Create a random intercept model
RATS_ref <- lmer(Weight ~ Time + Group + (1 | ID), data = ratsL, REML = FALSE)

# create a random intercept and random slope model
RATS_ref1 <- lmer(Weight ~ Time + Group + (Time | ID), data = ratsL, REML = FALSE)

# print a summary of the model
summary(RATS_ref1)

# perform an ANOVA test on the two models
anova(RATS_ref1, RATS_ref)

# create a random intercept and random slope model
RATS_ref2 <- lmer(Weight ~ Time * Group + (Time | ID), data = ratsL, REML = FALSE)

# print a summary of the model
summary(RATS_ref2)

# perform an ANOVA test on the two models
anova(RATS_ref2, RATS_ref1)

# draw the plot of RATSL
ggplot(ratsL, aes(x = Time, y = Weight, group = ID)) +
  geom_line(aes(linetype = Group)) +
  scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 20)) +
  scale_y_continuous(name = "Observed weight (grams)") +
  theme(legend.position = "top")

# Create a vector of the fitted values
Fitted <- fitted(RATS_ref2)

# Create a new column fitted to RATSL
RATSL <- ratsL %>%
  mutate(Fitted)

# draw the plot of RATSL
ggplot(RATSL, aes(x = Time, y = Fitted, group = ID)) +
  geom_line(aes(linetype = Group)) +
  scale_x_continuous(name = "Time (days)", breaks = seq(0, 60, 20)) +
  scale_y_continuous(name = "Fitted weight (grams)") +
  theme(legend.position = "top")
```

The RATS_ref model assumes independence of the repeated measures of weight, and this assumption is highly unlikely. In further analysis the weight is considered dependant to previous weight. 

The estimated regression parameters for the interaction indicate that the growth rate slopes are considerably higher for rats  in  group  2  than  for  rats  in  group  1  but  less  so  when comparing  group  3 rats  with  those  in  group  1. 

We  can  find  the  fitted  values  from  the  interaction  model  and  plot  the
fitted growth rates for each rat.  This graphic underlines how well the interaction model fits
the observed data
